Towards Verifying the Bitcoin-S Library

### 2019-06-26 paper/paper.tex
more intro

@@

What will turn out to be more relevant are the following Scala
features which Stainless does not support, such as: (concrete) class
definitions, inheritance by objects, abstract type members, variance
annotations, upper and lower type bounds in generics and private inner
classes.

\todo{anna: stainless unterstützt anscheinend type bounds prinzipiell
  (nur nicht selbstreferenzierende)? siehe B.4, stainless issue 519}

In addition, Stainless has its own library of some core data types and
functions which need to be mapped correctly to functions inside of the
SMT solver that Stainless ultimately relies on. Those data types in
general do not have all the methods of the Scala data types. We will
see an example of a method of the BigInt type that Scala provides but
Stainless does not.  \todo{question viktor: where is BigInt??}

\section{Trying to Verify the No-Inflation Property}

Thus, we try to verify the property, that a non-coinbase transaction
cannot generate new coins.  Let's name it the \nameref{property_1}.

As we can see later in chapter \ref{chap:verify_check}, we must
rewrite a huge part of the code implementing this property.  This
reimplementation into Pure Scala needs a lot of time.  So, we adjust
the plan and verify another functionality of Bitcoin-S.  Nevertheless,
during the analysis of the \nameref{property_1} we look at a bug in
Bitcoin-S found during this work and see the code changes for the
bugfix in section \ref{sec:bugfix}.

This chapter describes the part of Bitcoin-S needed to verify the
\nameref{property_1} described before.  We are going to create a
transaction and show the relevant parts of the method
\texttt{checkTransaction}, where transactions are checked against some
properties.  Then, we will see the bug in Bitcoin-S found during this
work and its fix.  In the end we see why the \nameref{property_1}
needed to be changed to the \nameref{property_2}.

\paragraph{Creating a Transaction}

Some code in this section is copied or adapted from the Bitcoin-S-Core
transaction builder example \cite{BitcoinSCore:txbuilderexample}.
Bitcoin-S-Core has a bitcoin transaction builder class with the
following constructor:
\begin{lstlisting}[style=scala]
  BitcoinTxBuilder(
    destinations: Seq[TransactionOutput], 
    utxos: BitcoinTxBuilder.UTXOMap,      
    feeRate: FeeUnit,                     
    changeSPK: ScriptPubKey,              
    network: BitcoinNetwork               
  ): Future[BitcoinTxBuilder]
\end{lstlisting}

@@

Second, we need destinations to spend the bitcoins to.
For the sake of convenience we create only one.
\begin{lstlisting}[style=scala]
  val destinationAmount = Satoshis(Int64(5000))

@@

\section{Towards Verifying the Addition-with-Zero Property}

>>>>

% to have bookmarks in pdf
% remove for final version
\usepackage[open,openlevel=5]{bookmark}
\setcounter{tocdepth}{5}

@@

What will turn out to be more relevant are the following Scala
features which Stainless does not support, such as: (concrete) class
definitions, inheritance by objects, abstract type members, variance
annotations and private inner classes.

In addition, Stainless has its own library of some core data types and
functions which need to be mapped correctly to functions inside of the
SMT solver that Stainless ultimately relies on. Those data types in
general do not have all the methods of the Scala data types. For
example, the BigInt type in Scala has a methods for bitwise operations
while the BigInt type in Stainless does not.

\paragraph{Outline.} In the next section we try to verify the property
that a regular (non-coinbase) transaction cannot generate new
coins. We call it the \emph{no-inflation property}. Trying to verify
it, we uncover and fix a bug in the bitcoin-s library. We find that
there is too much code involved that lies outside of the supported
fragment to currently make this verification feasible. Instead, in
Section~3 we turn to a simpler property to verify. The simplest
possible property we could think of was the fact that adding zero
satoshis to a given amount of satoshis yields the given amount of
satoshis. We call it the \emph{addition-with-zero property}. Here as
well we see that a significant part of the code lies outside of the
supported fragment. However, after a series of equivalent
transformations we arrive at code that we successfully
verify. \todo{conclusions: what now?}

\section{The No-Inflation Property}

\paragraph{Creating a Transaction} Some code in this section is copied or adapted from the Bitcoin-S-Core
transaction builder example \cite{BitcoinSCore:txbuilderexample}.
Bitcoin-S-Core has a bitcoin transaction builder class with the
following constructor:
\begin{lstlisting}[style=scala]
  BitcoinTxBuilder(
    destinations: Seq[TransactionOutput], 
    utxos: BitcoinTxBuilder.UTXOMap,      
    feeRate: FeeUnit,                     
    changeSPK: ScriptPubKey,              
    network: BitcoinNetwork               
  ): Future[BitcoinTxBuilder]
\end{lstlisting}

@@

Second, we need destinations to spend the bitcoins to.
For the sake of convenience we create only one.

\begin{lstlisting}[style=scala]
  val destinationAmount = Satoshis(Int64(5000))

@@

\section{The Addition-with-Zero Property}
